---
title: "Pre-processing step for Darwin Core mapping"
subtitle: "For: Inventory of alien species in Europe (DAISIE)"
author:
- Lien Reyserhove
- David Roy
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
    number_sections: yes
    toc: yes
    toc_depth: 3
    toc_float: yes
#  pdf_document:
#    df_print: kable
#    number_sections: yes
#    toc: yes
#    toc_depth: 3
---

This file describes the steps required to reorganize the input datasets. The goal of this reorganization is to improve the readability of the mapping scripts later.

# Setup 

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

Load libraries:

```{r}
library(tidyverse)      # To do data science
library(magrittr)       # To use %<>% pipes
library(janitor)        # To clean input data
library(readxl)         # To read Excel files
```

# Read data

1. Define data types

```{r}
col_types <- cols(.default = col_character())
```

2. Read literature references

```{r}
literature_references <- read_delim(
  file = "../data/raw/input_literature_references.csv", 
  delim = ",",
  col_types = col_types)
```

3. Read taxon data

```{r}
taxon <- read_delim(
  file = "../data/raw/input_taxon.csv", 
  delim = ",",
  col_types = col_types) 
```

4. Read distribution data:

```{r}
distribution <- read_delim(
  file = "../data/raw/input_distribution.csv", 
  delim = ",",
  col_types = col_types) 
```

# Compile reference data

The file `literature_references` includes all sources used to compile the DAISIE checklist.

1. Inspect `literature_references`:

```{r}
literature_references %>% head(n=10)
```

Each source can be identified by its `sourceid` and is composed of:
- a citation (`shortref`) and/or
- a full references (`longref`) and/or
- an url (`url`)

Most input files use the field `sourceid` to link with the sources included in `literature_references`. In the Darwin Core mapping process, we want to replace the `sourceid` with the full reference to each source. To accomplish this, we need to compile the full reference from:
- `shortref` and/or
- `longref` and/or
- `url`

2. Inspect longref:

```{r, echo=FALSE}
literature_references %>% 
  group_by(longref) %>% 
  summarize(records = n()) %>%
  arrange(longref)
```

Cleaning the conent of this field is outside scope op this mapping). However, several `longref` values start with a number (e.g. `1`, `1, 28`). These can easily be removed.

3. Remove numbers from `longref`

```{r, echo=FALSE}
literature_references %<>% mutate(longref = str_replace(longref, "^[0-9]", NA_character_))
```

4. Verify if numbers are removed (should be `TRUE`)

```{r}
unique(literature_references %>%  str_detect("^[0-9]")) == FALSE
```

5. Inspect `shortref`

```{r}
literature_references %>% 
  group_by(shortref) %>% 
  summarize(records = n()) %>%
  arrange(shortref)
```

Cleaning of this fiels is outside the scope of this mapping.

6. Inspect `url`

```{r}
literature_references %>% 
  group_by(url) %>% 
  summarize(records = n()) %>%
  arrange(url)
```

This information is, in most cases, wrong: `url` often refers to scientific names rather than to a real url. We remove the scientific names from `url` and keep real url values only.

7. Remove non-valid url values

```{r}
literature_references %<>%  mutate(url = if_else(str_detect(url, pattern = "^http"), url, NA_character_))
```

8. Verify cleaning step

```{r}
literature_references %>% 
  group_by(url) %>% 
  summarize(records = n()) %>%
  arrange(url)
```

8. Show mapping of full reference from `shortref`, `longref` and `url`

For most records, one or more of the abovementioned fields are empty. We need to know which (and how many) meaningfull combinations of these fields occur in `literature_references`: 

```{r}
overview_references <- arrange(
  bind_cols(
        data.frame(matrix(c(
                            "x","x","",          
                            "x","","",          
                            "x", "x", "x",   
                            "x","", "x",
                            
                            "","","",                            
                            "", "x","",
                            "","", "x",
                            "", "x", "x"),
                       nrow = 8, byrow = TRUE, 
                       dimnames = list(c(1:8), c("longref", "shortref", "url"))
                       )),
        data.frame(bind_rows(

          # "x","x",""  
          summarize(filter(literature_references,
                           !is.na(longref) &
                             !is.na(shortref) &
                             is.na(url)),
                     records = n()),
          
          # "x","",""
           summarize(filter(literature_references,
                           !is.na(longref) &
                             is.na(shortref) &
                             is.na(url)), 
                     records = n()),
          
          # "x", "x", "x",   
           summarize(filter(literature_references,
                           !is.na(longref) &
                             !is.na(shortref) &
                             !is.na(url)),
                    records = n()),
        
          # "x","", "x",
           summarize(filter(literature_references,
                           !is.na(longref) &
                             is.na(shortref) &
                             !is.na(url)),
                    records = n()),          
          # "","","",        
           summarize(filter(literature_references, 
                           is.na(longref) & 
                           is.na(shortref) & 
                           is.na(url)),
                    records = n()),
                              
          # "", "x",""
          summarize(filter(literature_references, 
                           is.na(longref) &
                             !is.na(shortref) &
                             is.na(url)),
                    records = n()),
  
          # "","", "x"
          summarize(filter(literature_references,
                            is.na(longref) &
                            is.na(shortref) &
                           !is.na(url)),
                    records = n()),  
          
          # "", "x", "x"                
          summarize(filter(literature_references,
                           is.na(longref) &
                             !is.na(shortref) &
                             !is.na(url)),
                    records = n())                    
                    
        ))
  )
)


overview_references
```

We use the following rules to generate `source`:

- If a longref is provided, never use a shortref. Url is integrated where provided
- If no longref is provided, use a shortref and/or Url:

```{r}
overview_references %>% mutate(source = case_when(
   
# Longref provided:
  longref == "x" & shortref == "x" & url == ""  ~ "longref",
  longref == "x" & shortref == ""  & url == ""  ~ "longref",
  longref == "x" & shortref == "x" & url == "x" ~ "longref + url",  
  longref == "x" & shortref == "" & url  == "x" ~ "longref + url",

# No longref provided:
  longref == ""  & shortref == ""  & url == ""  ~ "",
  longref == ""  & shortref == "x" & url == ""  ~ "shortref",
  longref == ""  & shortref == ""  & url == "x" ~ "url",
  longref == ""  & shortref == "x" & url == "x" ~ "shortref + url"))
```

9. Generate full reference (`source`)

```{r}
literature_references %<>% mutate(source = case_when(
  
# Longref provided:
  !is.na(longref) & !is.na(shortref) & is.na(url) ~ paste(longref),
  !is.na(longref) & is.na(shortref)  & is.na(url) ~ paste(longref),  
  !is.na(longref) & !is.na(shortref) & !is.na(url) ~ paste(longref, paste0("(",url,")"), sep = ""),  
  !is.na(longref) & is.na(shortref)  & !is.na(url) ~ paste(longref, paste0("(",url,")"), sep = ""),  
  
# No longref provided:
  is.na(longref)  & is.na(shortref)  & is.na(url) ~ "",
  is.na(longref)  & !is.na(shortref) & is.na(url) ~ paste(shortref),
  is.na(longref)  & is.na(shortref)  & !is.na(url) ~ paste(url),  
  is.na(longref)  & !is.na(shortref) & !is.na(url) ~ paste(shortref, paste0("(",url,")"), sep = "")
))  
```

We can now use `literature_references` to add the full reference to the generated Darwin Core files (except the distribution extension, see further), using `sourceid` as the link.

# Generate reference dataset for distribution extension 

The distribution extension has no `sourceid` field to link with the full reference in `literature_reference`.
In the case of the distribution extension, a specific reference can only be linked to a record using a combination between:
- `id_sp_region`
- `field_name`

We save this information in a separate dataframe `distribution_sources` here and export it to use it later to generate the full reference for the distribution extension.

1. Generate `distribution_sources

```{r}
distribution_sources <- 
  literature_references %>% 
    filter(!is.na(id_sp_region)) %>% 
    select(id_sp_region, field_name, sourceid, source)
```

2. Export as `distribution_sources`

```{r}
write_csv(distribution_sources, "../data/interim/distribution_sources.csv", na = "")
```

## Generate interim_literature_references

Wemove duplicated sources in `literature_references` (duplicates are only needed for the mapping of the distribution extension) and export them as interim file.

1. Remove duplicates:

```{r}
interim_literature_references <-
  literature_references %<>% distinct(sourceid, .keep_all = TRUE)
```

2. Export as `interim_literature_references`:

```{r}
write_csv(interim_literature_references, "../data/interim/interim_literature_references.csv", na = "")
```

# Generate additional files for description extension

`input_taxon` has one field that can't be used for mapping of the Taxon Core:
 - `ecofunct_group`

We will integrate this information in the description extension. 

1. Save information on ecofunctional groups (create `ecofunct_group`):

```{r}
ecofunctional_group <- 
  taxon %>% 
    select(idspecies, ecofunct_group, sourceid) %>% 
    distinct(idspecies, ecofunct_group, .keep_all = TRUE)
```

2. Remove records for which `ecofunct_group` is empty:

```{r}
ecofunctional_group %<>% filter(!is.na(ecofunct_group))
```

3. Export as `ecofunctional_group`

```{r}
write_csv(ecofunctional_group, "../data/interim/ecofunctional_group.csv", na = "")
```

