---
title: "Darwin Core mapping"
subtitle: "For: Inventory of alien species in Europe (DAISIE)"
author:
- Lien Reyserhove
- Peter Desmet
- David Roy
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
    number_sections: yes
    toc: yes
    toc_depth: 3
    toc_float: yes
#  pdf_document:
#    df_print: kable
#    number_sections: yes
#    toc: yes
#    toc_depth: 3
---

# Setup 

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

Load libraries:

```{r}
library(tidyverse)      # To do data science
library(magrittr)       # To use %<>% pipes
library(here)           # To find files
library(janitor)        # To clean input data
library(readxl)         # To read Excel files
library(digest)         # To generate hashes
library(rgbif)          # To use GBIF services
```

# Read source data

```{r}
taxon <- read.csv(here("data", "raw", "input_taxon.csv")) 
vernacular_names <- read.csv(here("data", "raw", "input_vernacular_names.csv"))
distribution <- read.csv(here("data", "raw", "input_distribution.csv"))

```

# Taxon core

Preview data:

```{r}
taxon %>% head(n = 5)
```

## Pre-processing

### Tidy data

Clean data somewhat:

```{r}
taxon <- 
  taxon %>% 
    remove_empty("rows") %>% 
    clean_names()
  
```

## Term mapping

Map the data to [Darwin Core Taxon](http://rs.gbif.org/core/dwc_taxon_2015-04-24.xml).

Start with record-level terms which contain metadata about the dataset (which is generally the same for all records).

### language

```{r}
taxon %<>% mutate(dwc_language = "en") 
```

### license

```{r}
taxon %<>% mutate(dwc_license = "http://creativecommons.org/licenses/by/4.0/legalcode") 
```

### rightsHolder

```{r}
taxon %<>% mutate(dwc_rightsHolder = "Centre for Ecology and Hydrology")
```

### reference


### datasetID

```{r}
taxon %<>% mutate(dwc_datasetID = "") 
```

### institutionCode

```{r}
taxon %<>% mutate(dwc_institutionCode = "CEH") 
```

### datasetName

```{r}
taxon %<>% mutate(dwc_datasetName = "Inventory of alien species in Europe (DAISIE)") 
```

The following terms contain information about the taxon:

### taxonID

```{r}
taxon %<>% mutate(dwc_taxonID = idspecies)
```

### scientificName

The information in `scientificName` is contained in several fields: `sp_genus`, `sp_species`, `sp_authority`, `sp_subtaxon` and `sp_subtaxon_authority`. We paste this information together to generate the field `dwc_scientificName`:

```{r}
taxon %<>% mutate(dwc_scientificName = paste(sp_genus, sp_species, sp_authority, sp_subtaxon, sp_subtaxon_authority, sep = " "))
```

Use the [GBIF nameparser](https://www.gbif.org/tools/name-parser) to retrieve nomenclatural information for the scientific names in the checklist:

```{r}
parsed_names <- taxon %>%
  distinct(dwc_scientificName) %>%
  pull() %>% # Create vector from dataframe
  parsenames() # An rgbif function
```

### kingdom

### phylum

```{r}
taxon %<>% mutate(dwc_phylum = sp_phylum)
```

### class

```{r}
taxon %<>% mutate(dwc_class = sp_class) 
```

### order

```{r}
taxon %<>% mutate(dwc_order = sp_ordo) 
```

### family

```{r}
taxon %<>% mutate(dwc_family = sp_family) 
```

### genus

```{r}
taxon %<>% mutate(dwc_genus = sp_genus) 
```

### specificEpithet

```{r}
taxon %<>% mutate(dwc_specificEpithet = sp_species) 
```

### infraspecificEpithet

```{r}
taxon %<>% mutate(dwc_infraspecificEpithet = sp_subtaxon) 
```

### taxonRank

### scientificNameAuthorship

### taxonRemarks

This field will include the information on ecofunctional groups (contained in `eg_ecofunct_group`)

```{r}
taxon %<>% mutate(dwc_taxonRemarks = paste("Ecofunctional_group", "=", eg_ecofunct_group, sep = " ")) 
```

## Post-processing

Only keep the Darwin Core columns:

```{r}
taxon %<>% select(starts_with("dwc_"))
```

Drop the `dwc_` prefix:

```{r}
colnames(taxon) <- str_replace(colnames(taxon), "dwc_", "")
```

Preview data:

```{r}
taxon %>% head()
```

Save to CSV:

```{r}
write_csv(taxon, here("data", "processed", "taxon.csv"), na = "")
```

# Vernacular names extension

Preview data:

```{r}
vernacular_names %>% head(n = 5)
```

## Tidy data

Clean data somewhat:

```{r}
vernacular_names <- 
  vernacular_names %>% 
    remove_empty("rows") %>% 
    clean_names()
  
```

## Term mapping

### taxonID

```{r}
vernacular_names %<>% mutate(dwc_taxonID = idspecies) 
```

### vernacularName

```{r}
vernacular_names %<>% mutate(dwc_vernacularName = cn_name) 
```

### source

```{r}

```

### language

Display all unique language information:

```{r}
vernacular_names %>% 
  group_by(cn_language) %>% 
  summarize(records = n())
```

Map language information to [ISO 639-1 Language Codes](https://tools.gbif.org/dwca-validator/vocabulary.do?id=http://iso.org/639-1)

```{r}
vernacular_names %<>% mutate(dwc_language = recode(cn_language,
  "Belorussian" = "be",
  "Czech" = "cs",
  "Danish" = "da",
  "Dutch" = "nl",
  "English" = "en",
  "Estonian" = "et",
  "Faeroese" = "fo",
  "Finnish" = "fi",
  "French" = "fr",
  "German" = "de",
  "Greek" = "el",
  "Hebrew (western characters)" = "he",
  "Hebrew (Hebrew characters)" = "he",
  "Hungarian" = "hu",
  "Icelandic" = "is",
  "Italian" = "it",
  "Latvian" = "lv",
  "Lithuanian" = "lt",
  "Maltese" = "mt",
  "Norwegian" = "no",
  "Polish" = "pl",
  "Portuguese" = "pt",
  "Romanian" = "ro",
  "Spanish" = "es",
  "Swedish" = "sv",
  "Turkish" = "tr"
))
```

Show mapping:

```{r}
vernacular_names %>% 
  group_by(cn_language, dwc_language) %>% 
  summarize(records = n())
```

## Post-processing

Only keep the Darwin Core columns:

```{r}
vernacular_names %<>% select(starts_with("dwc_"))
```

Drop the `dwc_` prefix:

```{r}
colnames(vernacular_names) <- str_replace(colnames(vernacular_names), "dwc_", "")
```

Preview data:

```{r}
vernacular_names %>% head()
```

Save to CSV:

```{r}
write_csv(vernacular_names, here("data", "processed", "vernacular_names.csv"), na = "")
```

# Distribution extension

